---
title: "Data Description"
author: "Sohoon Youn"
date: "2023-04-13"
output: pdf_document
---
```{r echo = FALSE, include=FALSE}
library(mlbench)
library(dplyr)
library(randomForest)
library(locfit)
library(mgcv)
library(dplyr)
library(locfit)
library(dplyr)
library(splines)
library(lattice)
library(flexmix)
library(mgcv)
library(nlme)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(kableExtra)
library(knitr)
library(caret)
library(rpart.plot)
library(randomForestSRC)
library(glmnet)
```

```{r echo = FALSE}
load("final.Rdata")
```


# 1.Data Description
## 1-1. Variables

This dataset contains 37 features, including `los`, `age`, `gender`, `ethnicity`, `religion`, `insurance`     `admission_type`, `admission_location`, `first_hosp_stay`, `diagnosis`, `los_reg_trs`, `aniongap_min`,    `aniongap_max`, `bicarbonate_min`, `bicarbonate_max`, `chloride_min`, `chloride_max`, `height`, `weight`, `hematocrit_min`, `hematocrit_max`, `hemoglobin_min`, `hemoglobin_max`, `platelet_min`, `platelet_max`, `bilirubin_max`, `sodium_max`, `potassium_max`, `heartrate_min`, `heartrate_max`, and `heartrate_mean`. 


```{r echo = FALSE}
model = lm(los~.,data=dat)
#summary(model)
name = c("ethnicity(black)", "religion", "aniongap_min", "bicarbonate_min","bicarbonate_max", "chloride_min", "chloride_max", "height", "weight", "heartrate_mean")
p.val = c(0.023371,0.022832,0.000696,0.012708,0.000209,6.46e-13,7.10e-08, 2e-16,8.14e-14,3.91e-05)
pval=data.frame(name,p.val)
```

```{r echo = FALSE}
kable(pval) %>% 
  kable_styling(position = "center", latex_options = "HOLD_position")
```

A full model is built in order to identify the significant variables. From the summary of this full model, the variable `ethnicity(black)`, `religion`, `aniongap_min`, `bicarbonate_min`,`bicarbonate_max`, `chloride_min`, `chloride_max`, `height`, `weight`, `heartrate_mean` are significant based on each p-value. 

## 1-2. New variables 

```{r echo = FALSE}
dat$aniongap = (dat$aniongap_min + dat$aniongap_max)/2
dat$bicarbonate = (dat$bicarbonate_min + dat$bicarbonate_max)/2
dat$chloride = (dat$chloride_min + dat$chloride_max)/2
dat$hematocrit = (dat$hematocrit_min + dat$hematocrit_max)/2
dat$hemoglobin = (dat$hemoglobin_min + dat$hemoglobin_max)/2
dat$platelet = (dat$platelet_min + dat$platelet_max)/2
```

The new features `aniongap`, `bicarbonate`, `chloride`, `hematocrit`, `hemoglobin`, and `platelet` are added to dataset since the given data has both max and min of those features. 


* `aniongap` :  An average of aniongap_min and aniongap_max

* `bicarbonate` :  An average of bicarbonate_min and bicarbonate_max

* `chloride` :  An average of chloride_min and chloride_max

* `hematocrit` :  An average of hematocrit_min and hematocrit_max

* `hemoglobin` :  An average of hemoglobin_min and hemoglobin_max

* `platelet` :  An average of platelet_min and platelet_max

## 1-3. Outliers
```{r echo = FALSE}
p1 = ggplot(data=dat, aes(x=weight, y=los)) + geom_point(alpha = 0.4,color=rgb(0.2,0.4,0.6,0.6)) + xlab('weight')+ylab('los')
p2 = ggplot(data=dat, aes(x=height, y=los)) + geom_point(alpha = 0.4,color=rgb(0.2,0.4,0.6,0.6)) + 
  xlab('height')+ylab('los')
p3 = ggplot(data=dat, aes(x=aniongap, y=los)) + geom_point(alpha = 0.4,color=rgb(0.2,0.4,0.6,0.6)) + 
  xlab('aniongap')+ylab('los')
p4 = ggplot(data=dat, aes(x=bicarbonate, y=los)) + geom_point(alpha = 0.4,color=rgb(0.2,0.4,0.6,0.6)) + 
  xlab('bicarbonate')+ylab('los')
p5 = ggplot(data=dat, aes(x=chloride, y=los)) + geom_point(alpha = 0.4,color=rgb(0.2,0.4,0.6,0.6)) + 
  xlab('chloride')+ylab('los')
p6 = ggplot(data=dat, aes(x=hematocrit, y=los)) + geom_point(alpha = 0.4,color=rgb(0.2,0.4,0.6,0.6)) + 
  xlab('hematocrit')+ylab('los')
p7 = ggplot(data=dat, aes(x=sodium_max, y=los)) + geom_point(alpha = 0.4,color=rgb(0.2,0.4,0.6,0.6)) + 
  xlab('sodium_max')+ylab('los')
p8 = ggplot(data=dat, aes(x=hemoglobin, y=los)) + geom_point(alpha = 0.4,color=rgb(0.2,0.4,0.6,0.6)) + 
  xlab('hemoglobin')+ylab('los')

ggarrange(p1,p2,p3,p4,ncol=2,nrow=2)
ggarrange(p5,p6,p7,p8,ncol=2,nrow=2)

```

* `height` was found to have a linear trend with `los`.

* `weight` was found to have a linear trend in certain weight divisions.

* `aniongap`, `bicarbonate`, `chloride`, `hematocrit`, `sodium_max`, and `hemoglobin` were hardly found to have a linear trend with `los`.

* Outliers are found from `dat` based on each scatter plot of `weight`, `height`, `aniongap`, `bicarbonate`, `chloride`, `hematocrit`, `sodium_max`, `heartrate_min`.


## 1-4. Variable Importance - Random Forest
```{r echo = FALSE}
fit = randomForest(los ~ ., 
                   data = dat, importance=TRUE)

imp=data.frame(importance(fit))
ord <- order(imp$X.IncMSE, decreasing=TRUE)
imp=imp[ord, ]
imp2 = imp[1:10,]
```
```{r echo = FALSE}
# Variable Importance
a=c("height", "weight", "heartrate_mean", "heartrate_min","admission_location", "hematocrit_min","hematocrit_max", "hematocrit","chloride_max", "chloride_min")
par(mfrow=c(1,2))
barplot(imp2$X.IncMSE,col="#69b3a2",ylim=c(0,50),ylab="X.IncMSE",names.arg=a,cex.names = 0.5,space=c(0.3,0.3,0.3,0.3,0.3),las=2,cex.axis=0.5)
barplot(imp2$IncNodePurity,col=rgb(0.8,0.1,0.1,0.6),ylim=c(0,350000),ylab="IncNodePurity",names.arg=a,cex.names = 0.5,space=c(0.3,0.3,0.3,0.3,0.3),las=2,cex.axis=0.5)
```
The importance of each variable is measured by the percentage increase in the mean squared error (MSE) when that variable is removed from the model. The higher the percentage increase in MSE, the more important the variable is to the model.

The two plots represent the IncreasedMSE and IncreasedNodepurity of each variable. According to the two plots, `height`, `weight`, `heartrate_mean`, are highly significant, while `admission_location`, `hematocrit`, `chloride_max`, `hematocrit_min`, `aniongap_min`, `hemoglobin_max` are weakly significant.


## 1-5. Interaction terms
```{r echo = FALSE}
cor_df = data.frame()
name = c('weight * height','height * heartrate_mean','chloride_min * sodium_max',
         'hematocrit * hemoglobin')
cor_ = c(cor(dat$weight,dat$height),cor(dat$height,dat$heartrate_mean),
         cor(dat$chloride_min,dat$sodium_max),cor(dat$hematocrit,dat$hemoglobin))
cor_df = data.frame(name,cor_)
colnames(cor_df) = c('Interactions','Correlation')

p9 = ggplot(data=dat, aes(x=height, y=weight)) + geom_point(alpha = 0.4,color=rgb(0.2,0.4,0.6,0.6)) + 
  xlab('height')+ylab('weight')
p10 = ggplot(data=dat, aes(x=height, y=heartrate_mean)) + geom_point(alpha = 0.4,color=rgb(0.2,0.4,0.6,0.6)) + 
  xlab('height')+ylab('heartrate_mean')
p11 = ggplot(data=dat, aes(x=chloride_min, y=sodium_max)) + geom_point(alpha = 0.4,color=rgb(0.2,0.4,0.6,0.6)) + 
  xlab('chloride_min')+ylab('sodium_max')
p12 = ggplot(data=dat, aes(x=hematocrit, y=hemoglobin)) + geom_point(alpha = 0.4,color=rgb(0.2,0.4,0.6,0.6)) + 
  xlab('hematocrit')+ylab('hemoglobin')

ggarrange(p9,p10,p11,p12,ncol=2,nrow=2)
```


```{r echo = FALSE}
kable(cor_df) %>% 
  kable_styling(position = "center", latex_options = "HOLD_position")
```

Correlations between variables are calculated, and scatter plots are drawn. According to the table, correlations between `weight&height`, `height&heartrate_mean`, `sodium_max&chloride_min`, and `hematocrit&hemoglobin` seem to be significant as the absolute values of correlations are significantly large. Based on scatter plots, `sodium_max&chloride_min` and `hematocrit&hemoglobin` were found to have a strong linear trend, while `height&heartrate_mean` has a relatively weak trend. 

Therefore, they are considered to be included in the following models. 

## 1-6. Histogram of los
```{r echo = FALSE}
par(mfrow=c(1,1))
hist(dat$los,col='skyblue',ylim=c(0,700),xlim=c(0,200),main='Histogram of los',cex.lab = 0.7,cex.main = 1,cex.axis = 0.7,xlab='los')
```

The feature `los` will be predicted by the following two models. As this project aims to predict `los` with high accuracy, it is crucial to examine the frequency of this feature. Therefore, the frequency of `los` is plotted, and it is found to be highly concentrated when `los` is less than 30. 